---
title: "Activity Monitoring"
author: "Claudio"
date: "21/07/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and preprocessing the data

We stop the process if the csv file is not in the working directory.  
Storing data into a data frame, without addressing NA's
Storing data without NA in another data frame

```{r read}
file <- "activity.csv"
  
if (!file.exists(file)) {
        stop("activity.csv file not found")
}

activity <- read.csv(file)
NoNAActivity <- subset(activity, is.na(activity$steps) == F)
```

# What is mean total number of steps taken per day?

## Calculating the total number of steps taken each day
```{r totalsteps}
library(plyr)
stepsPerDay <- ddply(NoNAActivity, .(date), summarise, steps=sum(steps))
```

## Creating the histogram
```{r histogramsteps}
hist(stepsPerDay$steps, breaks = 20, main="Number of Steps a day", xlab="Total number of steps taken each day", ylab = "Number of Days", col="cyan")
```

## Calculate and report the mean and median total number of steps taken per day
1. Mean
```{r Mean}
mean(stepsPerDay$steps)
```

2. Median
```{r Median}
median(stepsPerDay$steps)
```

